---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MultistateModelsPaper

<!-- badges: start -->
<!-- badges: end -->

MultistateModelsPaper provides a package for simulating and fitting multistate models
especially semi-Markov models, to panel data and interval censored data. 
This is a wrapper to the Julia package MultistateModels.jl. 
This version of the package is a stable version used to accompany the paper
"Assessing treatment efficacy for interval censored endpoints using multistate semi-Markov models fit to multiple data streams". 




## Installation and Setup

First, ensure `Julia` (<span>&#8805;</span> version 1.10) is installed. Julia can be downloaded from https://julialang.org/downloads/. Julia must be added to the system PATH; instructions can be found here: https://julialang.org/downloads/platform/. After installing Julia, ensure that the `JuliaConnectoR` R package has been installed and loaded. This package depends on having R version <span>&#8805;</span> 3.2).

```{r, warning=F, message=F}
#install.packages("JuliaConnectoR")
library(JuliaConnectoR)
```

Once the `JuliaConnectoR` package is loaded, the following code can be run to ensure Julia is installed correctly:

```{r, class.source='fold-show'}
JuliaConnectoR::juliaSetupOk()
```

If `Julia` is properly installed, this will return `TRUE`. If `FALSE` is returned, it means that `Julia` has not been installed correctly. See the help page for the `JuliaConnectoR-package` for more details. 

`dplyr`, `ggplot2`, `knitr`, and `kableExtra` are also required to be installed:

```{r, eval=F}
install.packages("dplyr")
install.packages("ggplot2")
install.packages("knitr")
install.packages("kableExtra")
```

The necessary Julia functions can be installed by running the following code:

```{r, eval=FALSE}
if (JuliaConnectoR::juliaSetupOk()){
    JuliaConnectoR::juliaEval('
       import Pkg
       Pkg.add(url = "https://github.com/fintzij/MultistateModels.jl#biostatistics_manuscript_2024")
       Pkg.add("CSV")
       Pkg.add("DataFrames")
       Pkg.add("Random")')
  } else {
    stop("Julia setup incorrect.
         Ensure Julia version >= 1.10 is properly installed.")
  }
```


If there is an error, see the help page for the `JuliaConnectoR-package` for more details. 

You can install MultistateModelsPaper from
[GitHub](https://github.com/) with:

```{r, warning=FALSE, message=FALSE}
#install.packages("devtools")
devtools::install_github("ammateja/MultistateModelsPaper", quiet=TRUE, build_vignettes = TRUE)
library(MultistateModelsPaper)
```


The vignette `illness-death` shows how to fit a basic 3-state illness-death model. 
In this simple setting, patients are healthy at the start of follow-
up. The model has three states – healthy, ill, and dead – and three transitions — healthy to ill, healthy to dead, and ill to dead. Disease recurrence is interval censored but the time
of death is exactly observed.

The vignette `Regeneron-simulation` shows how to simulate and analyze data with a similar structure to the Regeneron data used as a motivating example in the paper "Assessing treatment efficacy for interval censored endpoints using multistate semi-Markov models fit to multiple data streams". 

